Reduced plasticity and relative variance in physiological rates of ectotherm populations under climate change

### 2024-06-17 ms.qmd
More revisions for today. Mostly working on figures which are taking ages

#### *Multi-level Meta-analysis (MLMA) Models*
We first fit multi-level meta-analysis (MLMA) models (i.e., intercept-only models) for both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, that included study, species, and phylogeny as random effects to account for non-independence. We also included trait as a random effect to account for trait variation within the data. Our MLMA models allowed us to partition the variation in $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ among these key sources while accounting for total sampling variance in each. This allowed us to calculate the proportion of total heterogeneity [i.e., $I_{total}^2$; *sensu* @Nakagawa2012-oc; @Noble2022-ty] along with various $I^2$ metrics describing the proportion of variance explained by each random effect level [@Nakagawa2012-oc]. We also present 95% prediction intervals which describe the expected distribution of effects from future studies [@Nakagawa2021-ls; @Noble2022-ty].

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean (SD) = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean (SD) = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]`, n = `r acclim_time[3,4]`) than both freshwater (mean (SD) = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]`, n = `r acclim_time[1,4]`) and marine species (mean (SD) = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]`, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023-lz], which would make it more likely that terrestrial ectotherms would show lower post acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to $lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual variation in physiological rates across 10&deg;C differ depending on whether acute or acclimation responses are measured. If the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit a series of models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged we limited to trait categories for each habitat with six or more effect sizes.

@@

We explored the possibility for publication bias graphically, using funnel plots, and more formally by including in our meta-regression models sampling variance (or sampling standard error) [@Nakagawa2022-gy]. Funnel plot asymmetry may suggest a form of publication bias called the 'file-drawer' effect whereby low-powered studies are less likely to be published. To test whether sampling variance covaried with effect size we included it in a multi-level meta-regression model that accounted for all the random effects (study, species, trait) and fixed effects (acclimation time, type of effect, habitat, trait category and the interaction between habitat type and trait category).

@@

##################
  # Figure 1b plot
##################
col <- c("#0871B9", "#F3B40D", "#1BB908")

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 8, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

@@

  scale_fill_manual(values = c(col)) + ylim(-0.8, 2) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.7, size = 5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
        axis.text = element_text(size = 15), legend.title = element_text(size = 15),
                               legend.text = element_text(size = 15),
                               title = element_text(size = 24))

@@

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 8, branch.size = 2, k.pos = "none") +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +

  annotate("text", x = x1, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 3), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(f_cvr_mean), digit = 2)), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x2, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 3), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x2-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(m_cvr_mean), digit = 2)), hjust = 0, size = 6) +
  annotate("text", x = x2-0.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x3, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x3-0.15, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = x3-0.15, y = -1.5, label = paste0("P = ", round(pmcmc(t_cvr_mean), digit = 3)), hjust = 0, size = 6) + scale_fill_manual(values = c(col)) + ylim(-2, 2) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 15),
                                       legend.title = element_text(size = 15),
                               legend.text = element_text(size = 15),
                               title = element_text(size = 24))

  ggsave(p2, filename = "./output/figures/fig2.png", width = 9.3, height = 8.8)

@@

Relative variance in physiological rates ($lnCVR_{Q_{10}}$) showed a decrease with increasing temperature across all habitat types, which was especially pronounced in terrestrial and marine ectotherms. Reductions in variance were driven by slower increases in the variance in physiological rates relative to the mean with scaling relationships between log standard deviation and log mean all being smaller than 1 and varying across habitats (See Supplementary Materials). Overall, there was a `r mean(t_cvr_mean_ef)`% (95% CI: `r quantile(t_cvr_mean_ef, 0.025)[1]` to `r quantile(t_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(t_cvr_mean)`) reduction in relative physiological rate variance for terrestrial ectotherms and a `r mean(m_cvr_mean_ef)`% (95% CI: `r quantile(m_cvr_mean_ef, 0.025)[1]` to `r quantile(m_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(m_cvr_mean)`) reduction in relative variation for marine ectotherms when temperatures increased by 10$^\circ$C. In contrast, freshwater ectotherms exhibited a smaller reduction in relative physiological rate variance at high temperatures (`r mean(f_cvr_mean_ef)`%, 95% CI: `r quantile(f_cvr_mean_ef, 0.025)[1]` to `r quantile(f_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(f_cvr_mean)`).

@@

# Look at the environment details across different habitats. Seems very similar. *** Check with ALEX ***
  lnCVRQ10_data_wild  %>% group_by(habitat) %>% summarise(mean(cv), mean(acf_all))

@@

  # Lets now use the whole posterior distribtion instead. This is a little hacky because the object is complicated

######## *** THE CODE BELOW NEEDS SOME NEW CHECKING AS ABOVE WAS MODIFIED! ****

@@

Using the ERA5 climate model, predictions of current global changes in physiological rate variance were generally conservative with our model explaining ~ 50% of the variation in the observed data ($R^2$ = `r round(mean(r2_2), 2)`, 95% CI: `r round(quantile(r2_2, 0.025)[1], 2)` to `r round(quantile(r2_2, 0.975)[1], 2)`). Across habitat types climate change is predicted to result in a `r abs(lapply(mean_perc_habitat, mean)$f)`% change in relative variance for freshwater systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[2]`%, $p_{MCMC}$ = `r pmcmc(mean_perc_habitat$f)`), a `r abs(lapply(mean_perc_habitat, mean)$m)`% reduction in relative variance for marine systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[2]`%, $p_{MCMC}$ = `r pmcmc(mean_perc_habitat$m)`), and a `r abs(lapply(mean_perc_habitat, mean)$t)`% reduction in relative variance for terrestrial systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[2]`%, $p_{MCMC}$ = `r p_value(pmcmc(mean_perc_habitat$t))`) under a RCP8.5 climate scenerio.

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in physiological rate variance (relative variance) for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenario relative to current climate conditions (% change).

@@

Understanding acclimation capacity and how variation in physiological rates changes across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011; @bush2016incorporating; @seebacher2023can; @Sanderson2023]. Here, we show that the relative variance ($lnCVR_{Q_{10}}$) in physiological rates of ectotherms decreased across terrestrial, marine and freshwater ecosystems as temperatures increase, with the decrease being particularly pronouced in terrestrial ectotherms (~28%). These effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results uncover an hitherto unrecognised dynamic where the benefits of acclimation may be counteracted by a decrease in trait variance.

### ***Consequences of reduced plasticity and variance in physiological rates across ectotherms***

@@

```{r, fig-cvrtrait}
#| label: fig-cvrtrait
#| fig-cap: "Acute and Acclimation $lnCVR_{Q_{10}}$ across traits for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. The x-axis is truncated for ease of visualisation."
#| echo: false

>>>>

#### *Multi-level Meta-analysis (MLMA) Models*
We first fit multi-level meta-analysis (MLMA) models (i.e., intercept-only models) for both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, that included study, species, trait type, and phylogeny as random effects to account for non-independence and identify sources of variability. Our MLMA models allowed us to partition the variation in $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ among these key sources while accounting for total sampling variance in each. This allowed us to calculate the proportion of total heterogeneity [i.e., $I_{total}^2$; *sensu* @Nakagawa2012-oc; @Noble2022-ty] along with various $I^2$ metrics describing the proportion of variance explained by each random effect level [@Nakagawa2012-oc]. We also present 95% prediction intervals which describe the expected distribution of effects from future studies [@Nakagawa2021-ls; @Noble2022-ty].

@@

After quantifying levels of heterogeneity, we fit a series of multi-level meta-regression (MLMR) models to test our key questions. In all models, we included the same random effects as we used in our MLMA models. Acclimation time varied from `r range(data_wide$acclim_period, na.rm = TRUE)[1]` to `r range(data_wide$acclim_period, na.rm = TRUE)[2]` days (mean $\pm$ SD = `r mean(data_wide$acclim_period, na.rm = TRUE)` $\pm$ `r sd(data_wide$acclim_period, na.rm = TRUE)` days), and terrestrial ectotherms were acclimated for a much shorter duration (mean $\pm$ SD = `r acclim_time[3,2]` $\pm$ `r acclim_time[3,3]` days, n = `r acclim_time[3,4]`) than both freshwater (mean $\pm$ SD = `r acclim_time[1,2]` $\pm$ `r acclim_time[1,3]` days, n = `r acclim_time[1,4]`) and marine species (mean $\pm$ SD = `r acclim_time[2,2]` $\pm$ `r acclim_time[2,3]` days, n = `r acclim_time[2,4]`). Rates of acclimation have been shown to be faster for many terrestrial groups compared to aquatic organisms [e.g., amphibians and reptiles have faster rates of acclimation than fishes; See @Einum2023-lz], which would make it more likely that terrestrial ectotherms would show lower post acclimation $lnRR_{Q_{10}}$. To control for these possible differences, acclimation time was mean-centered (mean = 0) and included in all our models. As such, all estimates can be interpreted as values for an average level of acclimation time (i.e., `r mean(data_wide$acclim_period, na.rm = TRUE)` days).

We first tested the degree to which acute and acclimation $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ effects varied by habitat type (i.e., terrestrial, freshwater, and marine ecosystems). Models included an interaction between effect type (i.e., acute or acclimation) and habitat. Reduced mean $lnRR_{Q_{10_{acclimation}}}$ relative to $lnRR_{Q_{10_{acute}}}$ indicates that acclimation to thermal environments results in partial compensation of physiological rates (i.e., phenotypic plasticity), whereas no differences between $lnRR_{Q_{10_{acute}}}$ and $lnRR_{Q_{10_{acclimation}}}$ suggests organisms are not capable of physiological plasticity [@Havird2020-mj; @Seebacher2015-wi]. In contrast, a difference in $lnCVR_{Q_{10_{acclimation}}}$ relative to $lnCVR_{Q_{10_{acute}}}$ would imply that changes in between individual variation in physiological rates across 10&deg;C differ depending on whether acute or acclimation responses are measured. For both $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$, if the interaction between effect type and habitat was not supported, then we fit a model that only contained additive effects of effect type and habitat. Following on from these models, we subset each habitat type and explored how mean $lnRR_{Q_{10}}$ and $lnCVR_{Q_{10}}$ changed across traits. Within each habitat (marine, freshwater, and terrestrial) we fit models that included an interaction between effect type (acute / acclimation) and trait category (as defined above). Variance in effects within trait categories appeared to vary depending on the trait type in question. Comparison of a model with and without heteroscedastic residual variance favored a model with heteroscedastic residual variance across trait categories ($\Delta AIC_{c}$; marine = 58, freshwater = 120, and terrestrial = 12). To ensure models converged, we did not include trait categories that had less than six effect sizes.

@@

We explored the possibility for publication bias graphically using funnel plots, and more formally by including sampling variance (or sampling standard error) in our meta-regression models  [@Nakagawa2022-gy]. Funnel plot asymmetry may suggest a form of publication bias called the 'file-drawer' effect whereby low-powered studies are less likely to be published. To test whether sampling variance covaried with effect size we included it in a multi-level meta-regression model that accounted for all the random effects (study, species, trait) and fixed effects (acclimation time, type of effect, habitat, trait category and the interaction between habitat type and trait category).

@@

##################
  # Figure 1b plot
##################
  viridis_cols <- c( "#21908CFF", "#440154FF","#228C22")

@@

p1.2 <- orchaRd::orchard_plot(model1.2, mod = "habitat", group  = "record_num", at = list(type = c("acute", "acclim")), by = "type", weights = "prop", xlab = TeX("$lnRR_{Q_{10}}$"), angle = 45, legend.pos = "top.left", trunk.size = 1, branch.size = 2) + scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) +
  theme(legend.direction = "vertical") +

@@

  scale_fill_manual(values = viridis_cols) + ylim(-0.5, 1.5) +  annotate("text", label= paste0("k = ", k_lnrr$k, " (", k_lnrr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.7, size = 5)

  p1.2 <- p1.2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 24),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24))

@@

  viridis_cols <- c( "#21908CFF", "#440154FF","#228C22")

  #sum_data_p2 <- mod_results(model2.2, mod = "habitat", group = "record_num", by = "type", at = list(type = c("acute", "acclim")))

  #ggplot(lnCVRQ10_data, aes(y = effect_size, x = habitat, colour = habitat, shape = type)) + ggbeeswarm::geom_quasirandom( method = "tukey", aes(color = habitat)) + 
   #scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-1.5, 1.5) + scale_colour_manual(values = viridis_cols) + 
   # theme_classic() + coord_flip()

  p2 <- orchard_plot(model2.2, xlab = TeX("$lnCVR_{Q_{10}}$"), by = "type", at = list(type = c("acute", "acclim")), mod = "habitat", group  = "record_num", weights = "prop", angle = 45, trunk.size = 1, branch.size = 2, k.pos = "none") +
              scale_x_discrete(labels = c("T" = "Terrestrial","M" = "Marine" ,"F" = "Freshwater")) + ylim(-1.5, 1.5) + scale_fill_manual(values = viridis_cols) +

  annotate("text", x = x1, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(f_cvr_mean), 3), "[", round(quantile(f_cvr_mean, 0.025), 2)," to ", round(quantile(f_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(f_cvr_mean), digit = 2))), hjust = 0, size = 6) + 
  annotate("text", x = x1-0.15, y = 1, 
           label =paste0(round(mean(exp(f_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x2, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(m_cvr_mean), 3), "[", round(quantile(m_cvr_mean, 0.025), 2)," to ", round(quantile(m_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x2-0.15, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(m_cvr_mean), digit = 2))), hjust = 0, size = 6) +
  annotate("text", x = x2-0.15, y = 1, 
           label =paste0(round(mean(exp(m_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) +

  annotate("text", x = x3, y = -2, 
           label =TeX(paste0("$\\mu = $", round(mean(t_cvr_mean), 3), "[", round(quantile(t_cvr_mean, 0.025), 2)," to ", round(quantile(t_cvr_mean, 0.975),2), "] ")), hjust = 0, size = 6) + 
  annotate("text", x = x3-0.15, y = 1, 
           label =paste0(round(mean(exp(t_cvr_mean)-1)*100, 2), "%"), hjust = 0, size = 6) + labs(shape = "Type") +
    annotate("text", x = x3-0.15, y = -1.5, label = TeX(paste0("$p_{MCMC} = $", round(pmcmc(t_cvr_mean), digit = 3))), hjust = 0, size = 6) + ylim(-1.5, 1.75) + 
    annotate("text", label= paste0("k = ", k_cvr$k, " (", k_cvr$sdy, ")"), x = c(0.8, 1.8, 2.8), y = 1.3, size = 5) + ylim (-2, 2)

  p2 <- p2 + theme_classic() + theme(axis.title  = element_text(size = 24),
                                       axis.text = element_text(size = 24),
                                       axis.text.y = element_text(angle = 45, size = 24),
                                       legend.title = element_text(size = 24),
                               legend.text = element_text(size = 24),
                               title = element_text(size = 24))

  ggsave(p2, filename = "./output/figures/fig2.pdf", width = 9.3, height = 8.8)

@@

Relative variance in physiological rates ($lnCVR_{Q_{10}}$) showed a decrease with increasing temperature across all habitat types, which was especially pronounced in terrestrial and marine ectotherms. There were no differences between $lnCVR_{Q_{10}}$ and $lnCVR_{Q_{10}}$. Reductions in variance were driven by slower increases in the variance in physiological rates relative to the mean with scaling relationships between log standard deviation and log mean all being smaller than 1 and varying across habitats (See Supplementary Materials). There was a `r mean(t_cvr_mean_ef)`% (95% CI: `r quantile(t_cvr_mean_ef, 0.025)[1]` to `r quantile(t_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(t_cvr_mean)`) reduction in relative physiological rate variance for terrestrial ectotherms and a `r mean(m_cvr_mean_ef)`% (95% CI: `r quantile(m_cvr_mean_ef, 0.025)[1]` to `r quantile(m_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(m_cvr_mean)`) reduction in relative variation for marine ectotherms when temperatures increased by 10$^\circ$C. In contrast, freshwater ectotherms exhibited a smaller reduction in relative physiological rate variance at high temperatures (`r mean(f_cvr_mean_ef)`%, 95% CI: `r quantile(f_cvr_mean_ef, 0.025)[1]` to `r quantile(f_cvr_mean_ef, 0.975)[1]`, $p_{MCMC}$ = `r pmcmc(f_cvr_mean)`).

@@

# Look at the environment details across different habitats. Seems very similar.
  lnCVRQ10_data_wild  %>% group_by(habitat) %>% summarise(mean(cv), mean(acf_all))

@@

@@

Using the ERA5 climate model, predictions of current global changes in physiological rate variance were generally conservative with our model explaining ~ 50% of the variation in the observed data ($R^2$ = `r round(mean(r2_2), 2)`, 95% CI: `r round(quantile(r2_2, 0.025)[1], 2)` to `r round(quantile(r2_2, 0.975)[1], 2)`). Climate change is predicted to result in a `r abs(lapply(mean_perc_habitat, mean)$f)`% reduction in relative variance for freshwater systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$f[2]`%, $p_{MCMC}$ = `r pmcmc(mean_perc_habitat$f)`), a `r abs(lapply(mean_perc_habitat, mean)$m)`% reduction in relative variance for marine systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$m[2]`%, $p_{MCMC}$ = `r pmcmc(mean_perc_habitat$m)`), and a `r abs(lapply(mean_perc_habitat, mean)$t)`% reduction in relative variance for terrestrial systems (95% CI: `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[1]` to `r lapply(mean_perc_habitat, function(x) quantile(x, c(0.025, 0.975)))$t[2]`%, $p_{MCMC}$ = `r p_value(pmcmc(mean_perc_habitat$t))`) under a RCP8.5 climate scenerio.

```{r, fig-fig6}
#| label: fig-fig6
#| echo: false
#| warning: false
#| message: false
#| fig-cap: Model predictions for the expected change in acclimation $lnCVR_{Q_{10}}$ across the globe for terrestrial, marine and freshwater ecthotherms. Predictions consider the uncertainty in random effects (i.e., species, phylogeny, study).  Predicted change in relative physiological rate variance for each population based on current temperatures (average from 2018-2022; A-C) as well as the expected change from current temperatures based on future temperature predictions (average from 2096-2100, D). Future climate predictions are the reduction in relative variance expected under a RCP8.5 climate scenario relative to current climate conditions (% change).

@@

Understanding acclimation capacity and how variation in physiological rates changes across populations and species is important for predicting the ecological and evolutionary consequences of climate change [@Chevin2017; @Chevin2010; @Bolnick2011; @bush2016incorporating; @seebacher2023can; @Sanderson2023]. Here, we show that the relative variance ($lnCVR_{Q_{10}}$) in physiological rates of ectotherms decreased across terrestrial, marine and freshwater ecosystems as temperatures increase, with the decrease being particularly pronounced in terrestrial ectotherms (~28%). These effects are expected to result in a decrease in the variability in physiological rates for marine and terrestrial ectotherms between ~4-13% under future climate change projections. Our results uncover an hitherto unrecognised dynamic where the benefits of acclimation may be counteracted by a decrease in trait variance.

### ***Implications of differing acclimation capacities and decreased relative variability in physiological rates among ectotherms***

@@

```{r, fig-cvrtrait}
#| label: fig-cvrtrait
#| fig-cap: "Acute and acclimation $lnCVR_{Q_{10}}$ across traits for A) marine, B) freshwater and C) terrestrial systems. k = total number of effect size estimates while the numbers in brackets indicate the number of species. Thick bars indicate 95% confidence intervals and thin bars indicate 95% prediction intervals. The x-axis is truncated for ease of visualisation."
#| echo: false
